Elegromagnetism Assignment Two

### 2017-10-12 em/latex/HW2.tex
EM

\title{Quantum Mechanics Assignment One}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem One}
\textbf{Boundaries:}

\begin{align*}
  \phi(\infty) &= 0\\
  \phi(x,0) &= 0\\
  \phi(0,y) &= 0\\
\end{align*}

Guess the solution is four charges in a rectangle. If this solution works, by the Uniqueness Theorem it is the only solution.

\begin{align*}
  \phi &= \phi_1 + \phi_2 + \phi_3 + \phi_4\\
\end{align*}

The potential due to the real point charge is:

\begin{align*}
  \phi_1(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y-d_2)^2}}\\
\end{align*}

We let the second and fourth-quadrant points have charge -q, so potential:

\begin{align*}
  \phi_2(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y-d_2)^2}}\\
  \phi_4(x,y) &= \frac{-q}{4\pi\epsilon\sqrt{(x-d_1)^2+(y+d_2)^2}}\\
\end{align*}

and the third-quadrant point have charge +q, so potential:

\begin{align*}
  \phi_3(x,y) &= \frac{q}{4\pi\epsilon\sqrt{(x+d_1)^2+(y+d_2)^2}}\\
\end{align*}

Thus the total potential is given by:

\begin{align*}
  \phi(x,y) &= \frac{q}{4\pi\epsilon}\Big(
  \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
  &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)
\end{align*}

This satisfies our two boundary conditions:

\begin{align*}
  \phi(0,y) &= \frac{q}{4\pi\epsilon}\Big(
  \left(d_1^2+(y-d_2)^2\right)^{-1/2} + \left(d_1^2+(y+d_2)^2\right)^{-1/2}
  -\left(d_1^2+(y-d_2)^2\right)^{-1/2} - \left(d_1^2+(y+d_2)^2\right)^{-1/2}\Big)
  = 0\\
\end{align*}

\begin{align*}
  \phi(x,0) &= \frac{q}{4\pi\epsilon}\Big(
  \left((x-d_1)^2+d_2^2\right)^{-1/2} + \left((x+d_1)^2+d_2^2\right)^{-1/2}
  -\left((x+d_1)^2+d_2^2\right)^{-1/2} - \left((x-d_1)^2+d_2^2\right)^{-1/2}\Big)
  = 0\\
\end{align*}

To find the force on the first-quadrant charge, we use Coulomb's Equation:

%% \begin{align*}
%%   F &= -q\left(\frac{d}{dx} + \frac{d}{dy}\right) \frac{q}{4\pi\epsilon}\Big(
%%   \left((x-d_1)^2+(y-d_2)^2\right)^{-1/2} + \left((x+d_1)^2+(y+d_2)^2\right)^{-1/2}\\
%%   &-\left((x+d_1)^2+(y-d_2)^2\right)^{-1/2} - \left((x-d_1)^2+(y+d_2)^2\right)^{-1/2}\Big)\\
%%   &= \frac{-q^2}{4\pi\epsilon} \Big(
%%   (x-d_1 + y-d_2)\left((x-d_1)^2+(y-d_2)^2\right)^{-3/2} - (x+d_1 + y+d_2)\left((x+d_1)^2+(y+d_2)^2\right)^{-3/2}\\
%%   &+ (x+d_1 + y-d_2)\left((x+d_1)^2+(y-d_2)^2\right)^{-3/2} - (x-d_1 + y+d_2)\left((x-d_1)^2+(y+d_2)^2\right)^{-3/2}\Big)\\
%%   F(d_1,d_2) &= \frac{-q^2}{4\pi\epsilon} \Big(
%%   - (2d_1 + 2d_2)\left(4d_1^2 + 4d_2^2\right)^{-3/2} + 2d_1\left(4d_1^2\right)^{-3/2} - 2d_2\left(4d_2^2\right)^{-3/2}\Big)\\
%%   &= \frac{q^2}{4\pi\epsilon} \left(\frac{(2d_1 + 2d_2)}{\sqrt{\left(4d_1^2 + 4d_2^2\right)}^3} - \frac{1}{4d_1^2} - \frac{1}{4d_2^2}\right)\\
%% \end{align*}

\begin{align*}
  \phi(\vec{r}) &= \frac{q}{\sqrt{2}4\pi\epsilon_0\sqrt{4d_1^2 + 4d_2^2}}\left(\hat{i} + \hat{j}\right) + \frac{-q}{4\pi\epsilon_0\sqrt{4d_1^2}}\hat{i} + \frac{-q}{4\pi\epsilon_0\sqrt{4d_2^2}}\hat{j}\\
  &= \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_1^2}}\right)\hat{i}
  + \frac{q}{4\pi\epsilon_0}\left(\frac{1}{\sqrt{8d_1^2 + 8d_2^2}} - \frac{1}{\sqrt{4d_2^2}}\right)\hat{j}
\end{align*}

If expanded, at large r, $\phi$ should look like a quadrupole scaling with $\frac{1}{r^3}$.\\

\section{Problem Two}
We want to show the following equality is always true:

\begin{align*}
  \int\rho_1\phi_2d^3\vec{r} &= \int\rho_2\phi_1d^3\vec{r}
\end{align*}

We first apply the Laplace Equation:

\begin{align*}
  -\epsilon_0\int\nabla^2\phi_1\phi_2d^3\vec{r} &= -\epsilon_0\int\nabla^2\phi_2\phi_1d^3\vec{r}
\end{align*}

We then expand this equation into its cartesian representation:

\begin{align*}
  \int\int\int\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_1) \phi_2 dxdydz &=
  \int\int\int\sum_{n=x,y,z}\frac{\partial^2}{\partial n^2} (\phi_2) \phi_1 dxdydz
\end{align*}

We then use Integration by Parts:

\begin{align*}
  \int\int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_1\phi_2\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_1\frac{\partial}{\partial n}\phi_2dn \right) dxdydz &=
  \int\int\int\sum_{n=x,y,z}\left(\frac{d}{dn}\phi_2\phi_1\Big|_{-\infty}^{\infty} - \int\frac{\partial}{\partial n}\phi_2\frac{\partial}{\partial n}\phi_1dn \right) dxdydz
\end{align*}

\section{Problem Three}
The solution to the full sphere in an external field $E_0$ was shown in class to be:

\begin{align*}
  \phi(r,\theta) &= \left(-E_0r + \frac{E_0a^3}{r^2}\right)\cos\theta
\end{align*}

given boundaries $\phi(r=a) = 0$ and $\phi(r\rightarrow\infty) = 0$. We notice that our boundaries for the half sphere problem are the same, plus one addition: $\phi(r<a, z=0) = 0$. The above solution satisfies this condition as well. Then by the Uniqueness Theorem this solution must also be valid for our half sphere.\\

The first term $-E_0r\cos\theta$ is the potential from the external field. The second term $\frac{E_0a^3\cos\theta}{r^2}$ is the term due
to the sphere. This term can be rewritten as the dipole term in the multipole expansion:\\

\begin{align*}
  \frac{E_0a\cos\theta}{r^2} &= \frac{1}{4\pi\epsilon_0}\frac{1}{r^2} \hat{r} \cdot \vec{p}\\
\end{align*}

where $\vec{p} = 4\pi\epsilon_0E_0 a^3 \hat{z}$. Thus the half sphere can be exactly expressed as a dipole.\\

>>>>

\title{Elegromagnetism Assignment Two}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\newcommand{\bra}[1]{\big<#1\big|}
\newcommand{\ket}[1]{\big|#1\big>}
\newcommand{\braket}[2]{\big<#1\big|#2\big>}

@@

\section{Problem One}
To find the bound charges on this cross section, we use the following equations:

\begin{align*}
  \rho_b &= -\nabla \vec{P} = 0\hspace{2cm} \mbox{P is constant}\\
  \sigma_b &= \vec{P} \cdot \vec{n}\\
\end{align*}

The top two faces each make a $60^\circ$ angle with $\vec{P}$, and the bottom one is antiparallel. Thus we see:

\begin{align*}
  \sigma_{b,top} &= \vec{P} \cdot \hat{n} = P\cos\left(60^\circ\right) = \frac P2\\
  \sigma_{b,bot} &= \vec{P} \cdot \hat{n} = P\cos\left(180^\circ\right) = -P\\
\end{align*}

To find total charge we just sum these densities over the surface area of the prism. We notice that there are two top faces and one
bottom face, thus the two surface charges cancel out and the net charge on the object is \textbf{zero}. We expect this for a dielectric with no free charge.\\

We then find the far field by treating the triangle as a point dipole extending infinitely through the z axis:

\begin{align*}
  \vec{E}(\vec{r}, \theta) &= \int_\infty^\infty \frac{\vec{P}\cdot\vec{r}}{\sqrt{z^2+r^2}}dz
  = P\cos(\theta)\int_\infty^\infty \frac{\sqrt{z^2+r^2}}{\sqrt{z^2+r^2}^3}dz
  = P\cos(\theta)\int_\infty^\infty \frac{1}{z^2+r^2}dz\\
  &= \frac{P\cos(\theta)\pi}{r}
\end{align*}

This result almost makes sense. Normally a dipole falls off by $\frac{1}{r^3}$, but this one is generated by an infinite line, so it
falls off lots slower.\\

\section{Problem Two}
The example solved in Griffiths Ex 4.7 can be generalized to a two-dielectrics-in-all-space system by letting $\epsilon_r = \epsilon_1 / \epsilon_2$:

\textbf{a.)}\\
\begin{align*}
	V_{in} &= -\frac{3E_0r\cos\theta}{\epsilon_r+2}\\
	E_{in} &= \frac{3E_0\cos\theta}{\epsilon_r+2}\hat{r} + \frac{-3E_0\sin\theta}{\epsilon_r+2}\hat{\theta}\\
        V_{out} &= -E_0r\cos\theta + \frac{(\epsilon_r-1)R^3E_0\cos\theta}{(\epsilon_r+2)r^2}\\
        E_{out} &= \left(E_0\cos\theta+\frac{2(\epsilon_r-1)R^3E_0\cos\theta}{(\epsilon_r+2)r^3}\right)\hat{r} + \left(-E_0\sin\theta+\frac{-(\epsilon_r-1)R^3E_0\sin\theta}{(\epsilon_r+2)r^3}\right)\hat{\theta}\\
        &= \left(\frac{E_0\cos\theta(3\epsilon_r-1)}{\epsilon_r+2}\right)\hat{r} + \left(\frac{-3E_0\sin\theta}{\epsilon_r+2}\right)\hat{\theta}\\
\end{align*}

\textbf{b.)}\\
Solving for free and bound surface charge:

\begin{align*}
  P &= \epsilon_1E\\
  \sigma_b &= P \cdot \hat{r} = \frac{3E_0\cos\theta}{\epsilon_r+2}\\
  \sigma_f &= \epsilon_1E_{in,r} - \epsilon_2E_{out,r} = \frac{-E_0\cos\theta\epsilon_2}{\epsilon_r+2}\\
  \sigma &= \frac{-E_0\cos\theta(3-\epsilon_2)}{\epsilon_r+2}\\
\end{align*}

This charge is positive at the top of the sphere and negative at the bottom, as expected for a bound charge.\\

\textbf{c.)}\\
Tangential E-fields are equal at $r=R$:

\begin{align*}
  E_{in,\theta} - E_{out,\theta} =& \frac{-3E_0\sin\theta}{\epsilon_r+2} - \frac{-3E_0\sin\theta}{\epsilon_r+2} = 0\\
\end{align*}

\textbf{d.)}
E-fields are curl-free. Because the dielectrics are linear, D is proportional to E, and so it seems like the difference should be zero. However, we see that:\\

\begin{align*}
  D_{out} - D_{in} &= \left(\epsilon_0-\epsilon_1\right)\frac{-3E_0\sin\theta}{\epsilon_r+2}
\end{align*}

The change in permittivity constants causes a change in curl.\\
